---
title: "Stochastic surplus production model in continuous time (SPiCT)"
author: "Marta Cousido-Rocha and Francisco Izquierdo"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: no
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: no
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
```

```{r,warning=FALSE,message=FALSE}
#remotes::install_github("tokami/spict/spict", ref="fix130")
library(spict)
library(readxl)
```

# Preparing data: Biomass catch time series {.tabset}


Load data:

```{r,warning=FALSE}
load("data_SPiCT new CPUE & original CPUE& Catch & LFD.RData")
attach(objs)
```

Sum catch data across fleets:

```{r,warning=FALSE}
library(plyr)
catch<-ddply(catch_by_fleet, .(year,seas), summarize,  cat=sum(catch))
```

Weight-length relationship:


```{r,warning=FALSE}
len=as.numeric(substr(colnames(LFD_by_fleet)[-(1:6)], 2, 4))
a=2.459*10^{-5}
b=2.9667
wei=a*len^b
```

LFDs sum across fleets:

```{r,warning=FALSE}
library(dplyr)

LFD<-LFD_by_fleet %>% group_by(Yr, Seas) %>% summarise(across(everything(), list(sum)))
LFD=LFD[,-(3:6)]
```



Combine catch (numbers), LFDs and weight to obtain catch (tons):

```{r,warning=FALSE}
l=dim(LFD)[1]

resul=as.data.frame(LFD[,1:2])
resul$cat=rep(0,l)
for (i in 1:l){
  
  tot=sum(LFD[i,-c(1,2)])
  
  lfd=(LFD[i,-c(1,2)]*(catch[i,3]/tot))
  
  resul[i,3]=sum(lfd*wei)
}

# Sum across year:

catch_year=ddply(resul, .(Yr), summarize,  cat=sum(cat))
```


List to save SPiCT trajectories:

```{r,warning=FALSE,include=FALSE}
Bio_spict=list()
F_spict=list()
```



# Run 1: Year time series (Bayesian CPUE) {.tabset}

```{r,warning=FALSE}
ind1=which(unique(resul$Yr)==1972)
ind2=which(unique(resul$Yr)==2015)
inp1 <- list(timeC = unique(resul$Yr)[ind1:ind2], obsC =catch_year$cat[ind1:ind2],
            timeI = list(cpue_year_besag$year+0.5),
            obsI = list(cpue_year_besag$obs_sum))

inp1=check.inp(inp1)

```

The data can be plotted using the command:

```{r,warning=FALSE,fig.height=5}
source("functions.R")
source("plotting.R")
plotspict.data.adp(inp1)

```


The model is fitted to data by running:

```{r,warning=FALSE}
res1 <- fit.spict(inp1)
```

The results are summarized using:


```{r,warning=FALSE}
capture.output(summary(res1))
plot(res1,CI = 0.9)
```

```{r,warning=FALSE}
par(mfcol=c(1,2))
plotspict.biomass(res1)
plotspict.f(res1)
```


```{r,warning=FALSE}
par(mfcol=c(1,1))
plotspict.production(res1)
```


**Checklist for the acceptance of a SPiCT assessment** 
  
  
```{r,warning=FALSE}
res1$opt$convergence 
```

```{r,warning=FALSE}
all(is.finite(res1$sd))  
```

```{r,warning=FALSE}
r1 <- calc.osa.resid(res1)
plotspict.diagnostic(r1,lag.max =8)
```

```{r,warning=FALSE}
r1<- res1
rep1=retro(r1, nretroyear=5)
plotspict.retro(rep1)
```


Confidence intervals of $F/F_{msy}$ and $B/B_{msy}$ estimates.

```{r,warning=FALSE}
fit=res1

get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9)
```

I also compute AIC value:
  
```{r,warning=FALSE}
get.AIC(res1)
```

Save biomass/F trends:

```{r}
bio1=get.par("logB", res1, exp = TRUE)[seq(1,length(inp1$time),by=16),2]
F1=get.par("logFs", res1, exp = TRUE)[seq(1,length(inp1$time),by=16),2]

Bio_spict$bio_year_besag=bio1
F_spict$F_year_besag=F1
```

# Run 2: Year time series (Original CPUE) {.tabset}

```{r,warning=FALSE}
ind1=which(unique(resul$Yr)==1972)
ind2=which(unique(resul$Yr)==2015)
inp2 <- list(timeC = unique(resul$Yr)[ind1:ind2], obsC =catch_year$cat[ind1:ind2],
            timeI = list(cpue_year_original$year+0.5),
            obsI = list(cpue_year_original$obs_sum))
inp2=check.inp(inp2)

```

The data can be plotted using the command:

```{r,warning=FALSE,fig.height=5}
plotspict.data.adp(inp2)

```


The model is fitted to data by running:
```{r,warning=FALSE}
res2 <- fit.spict(inp2)
```

The results are summarized using:


```{r,warning=FALSE}
capture.output(summary(res2))
plot(res2,CI = 0.9)
```

```{r,warning=FALSE,eval=FALSE, include=FALSE}
par(mfcol=c(1,2))
plotspict.biomass(res2)
plotspict.f(res2)
```


```{r,warning=FALSE}
par(mfcol=c(1,1))
plotspict.production(res2)
```




**Checklist for the acceptance of a SPiCT assessment** 
  
  
```{r,warning=FALSE}
res2$opt$convergence 
```

```{r,warning=FALSE}
all(is.finite(res2$sd))  
```

```{r,warning=FALSE}
r2 <- calc.osa.resid(res2)
plotspict.diagnostic(r2,lag.max =8)
```

```{r,warning=FALSE}
r2<- res2
rep1=retro(r2, nretroyear=5)
plotspict.retro(rep1)
```


Confidence intervals of $F/F_{msy}$ and $B/B_{msy}$ estimates.

```{r,warning=FALSE}
fit=res2

get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9)
```

I also compute AIC value:
  
```{r,warning=FALSE}
get.AIC(res2)
```

Save biomass/F trends:

```{r}
bio2=get.par("logB", res2, exp = TRUE)[seq(1,length(inp2$time),by=16),2]
F2=get.par("logFs", res2, exp = TRUE)[seq(1,length(inp2$time),by=16),2]

Bio_spict$bio_year_original=bio2
F_spict$F_year_original=F2
```


# Run 3: Seas catch and CPUE time series (Bayesian CPUE) {.tabset}

```{r,warning=FALSE}

ind=which(resul[,2]==1)
resul[ind,2]=0
ind1=which(resul[,2]==4)
resul[ind1,2]=0.25
ind2=which(resul[,2]==7)
resul[ind2,2]=0.5
ind3=which(resul[,2]==10)
resul[ind3,2]=0.75
seq_I=rep(c(0,0.25,0.5,0.75),length(unique(cpue_seas_besag$year)))
ind1=min(which(resul[,1]==1972))
ind2=max(which(resul[,1]==2015))
inp3 <- list(timeC = resul[ind1:ind2,1]+resul[ind1:ind2,2], obsC = resul[ind1:ind2,3],
            timeI = list(cpue_seas_besag$year+seq_I),
            obsI = list(cpue_seas_besag$obs))

inp3=check.inp(inp3)

```

The data can be plotted using the command:

```{r,warning=FALSE,fig.height=5}
plotspict.data.adp(inp3)

```


The model is fitted to data by running:
```{r,warning=FALSE}
res3 <- fit.spict(inp3)
```

The results are summarized using:


```{r,warning=FALSE}
capture.output(summary(res3))
plot(res3,CI = 0.9)
```

```{r,warning=FALSE,eval=FALSE, include=FALSE}
par(mfcol=c(1,2))
plotspict.biomass(res3)
plotspict.f(res3)
```


```{r,warning=FALSE}
par(mfcol=c(1,1))
plotspict.production(res3)
```



**Checklist for the acceptance of a SPiCT assessment** 
  
  
```{r,warning=FALSE}
res3$opt$convergence 
```

```{r,warning=FALSE}
all(is.finite(res3$sd))  
```

```{r,warning=FALSE}
r3 <- calc.osa.resid(res3)
plotspict.diagnostic(r3,lag.max =8)
```

```{r,warning=FALSE}
r3<- res3
rep3=retro(r3, nretroyear=5)
plotspict.retro(rep3)
```


Confidence intervals of $F/F_{msy}$ and $B/B_{msy}$ estimates.

```{r,warning=FALSE}
fit=res3

get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9)
```

I also compute AIC value:
  
```{r,warning=FALSE}
get.AIC(res3)
```

Save biomass/F trends:

```{r}
bio3=get.par("logB", res3, exp = TRUE)[seq(1,length(inp3$time),by=4),2]
F3=get.par("logFs", res3, exp = TRUE)[seq(1,length(inp3$time),by=4),2]

Bio_spict$bio_seas_besag=bio3
F_spict$F_seas_besag=F3
```


# Run 4: Seas catch and CPUE time series (Original CPUE) {.tabset}

```{r,warning=FALSE}

ind=which(resul[,2]==1)
resul[ind,2]=0
ind1=which(resul[,2]==4)
resul[ind1,2]=0.25
ind2=which(resul[,2]==7)
resul[ind2,2]=0.5
ind3=which(resul[,2]==10)
resul[ind3,2]=0.75
seq_I=rep(c(0,0.25,0.5,0.75),length(unique(cpue_seas_besag$year)))
ind1=min(which(resul[,1]==1972))
ind2=max(which(resul[,1]==2015))
inp4 <- list(timeC = resul[ind1:ind2,1]+resul[ind1:ind2,2], obsC = resul[ind1:ind2,3],
            timeI = list(cpue_seas_original$year+seq_I),
            obsI = list(cpue_seas_original$obs))

inp4=check.inp(inp4)

```

The data can be plotted using the command:

```{r,warning=FALSE,fig.height=5}
plotspict.data.adp(inp4)

```


The model is fitted to data by running:
```{r,warning=FALSE}
res4 <- fit.spict(inp4)
```

The results are summarized using:


```{r,warning=FALSE}
capture.output(summary(res4))
plot(res4,CI = 0.9)
```

```{r,warning=FALSE,eval=FALSE, include=FALSE}
par(mfcol=c(1,2))
plotspict.biomass(res4)
plotspict.f(res4)
```


```{r,warning=FALSE}
par(mfcol=c(1,1))
plotspict.production(res4)
```



**Checklist for the acceptance of a SPiCT assessment** 
  
  
```{r,warning=FALSE}
res4$opt$convergence 
```

```{r,warning=FALSE}
all(is.finite(res4$sd))  
```

```{r,warning=FALSE}
r4 <- calc.osa.resid(res4)
plotspict.diagnostic(r4,lag.max =8)
```

```{r,warning=FALSE}
r4<- res4
rep4=retro(r4, nretroyear=5)
plotspict.retro(rep4)
```


Confidence intervals of $F/F_{msy}$ and $B/B_{msy}$ estimates.

```{r,warning=FALSE}
fit=res4

get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9)
```

I also compute AIC value:
  
```{r,warning=FALSE}
get.AIC(res4)
```
Save biomass/F trends:

```{r}
bio4=get.par("logB", res4, exp = TRUE)[seq(1,length(inp4$time),by=4),2]
F4=get.par("logFs", res4, exp = TRUE)[seq(1,length(inp4$time),by=4),2]

Bio_spict$bio_seas_original=bio4
F_spict$F_seas_original=F4
```





```{r}

save(Bio_spict, file="Bio_spict.RData")
save(F_spict, file="F_spict.RData")
```

# Comparing to SS {.tabset}

```{r}
load("data_SPiCT new CPUE & original CPUE & Bio.RData")
bio=objs$bio
a=c(Bio_spict$bio_seas_original,Bio_spict$bio_seas_besag,Bio_spict$bio_year_original,
    Bio_spict$bio_year_besag,bio$bio)
ind1=min(which(bio$year==1972))
ind2=max(which(bio$year==2015))

par(mar = c(5, 4, 1.4, 0.2))
plot(bio$year[ind1:ind2],bio$bio[ind1:ind2], col="black", type="l",lwd=2, ylim=c(min(a),max(a)),xlab="",ylab="")

ind1=min(which(resul[,1]==1972))
ind2=max(which(resul[,1]==2015))
ind3=max(which(names(Bio_spict$bio_seas_original)=="2015.75"))
lines(as.numeric(names(Bio_spict$bio_seas_original)[1:ind3]), Bio_spict$bio_seas_original[1:ind3],col="green",lwd=2)


ind3=max(which(names(Bio_spict$bio_seas_besag)=="2015.75"))
lines(as.numeric(names(Bio_spict$bio_seas_besag)[1:ind3]),Bio_spict$bio_seas_besag[1:ind3], col="grey",lwd=2)

ind1=which(unique(resul$Yr)==1972)
ind2=which(unique(resul$Yr)==2015)
ind3=which(names(Bio_spict$bio_year_besag)=="2015")
lines(unique(resul$Yr)[ind1:ind2],Bio_spict$bio_year_original[1:ind3], col="blue",lwd=2)
lines(unique(resul$Yr)[ind1:ind2],Bio_spict$bio_year_besag[1:ind3], col="red",lwd=2)

legend("topleft", legend=c("SS","bio_seas_original", "bio_seas_besag","bio_year_original","bio_year_besag"),
       col=c("black","green","grey","blue","red"), cex=1, lty=1, bty='n',y.intersp=0.8)
```

